---
title: "Assignment4_Alketa"
author: "Alketa Ulaj"
date: "12th of March, 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Assignment 4
  
I. Data: The World Values Survey is an ongoing worldwide survey that polls the world population about perceptions of life, work, family, 
politics, etc. The most recent phase of the survey that polled 77,882 people from 57 countries estimates that 36.2% of the worlds 
population agrees with the statement "Men should have more right to a job than women." The survey also estimates that 13.8% of people 
have a university degree or higher and that 3.6% of people fit both criteria. 

Q1. Are agreeing with the statement "Men should have more right to a job than women" and having a university degree or higher disjoint events?

A1: The definition of disjoint events is that they cannot happen at the same time. According to the paragraph they are not disjoint events. 
    Due to the amount of people who agree with the statement and posses a uni degree.And nonetheless the probability is not equals 
    to 0 therefore we can come to the conclusion of them not being disjoint events.

Q2. Draw a Venn diagram summarizing the variables and their associated probabilities.

A2:
```{r}

library(VennDiagram) ## install package for Venn Diagram
library(grid) ## install grid package

venn_summary <- draw.pairwise.venn(36.2, 13.8, 3.6, c("Population agreeing", "People who have a degree"),  
scale = FALSE, 
fill = c("blue", "red"));
grid.draw(venn_summary)

```

Q3. What is the probability that a randomly drawn person has a university degree or higher or agrees with 
   the statement about men having more right to a job than women?

A3: 0.464 
```{r}

## p(A) = 0.362 p(B) = 0.138 p(A&B) = 0.036
## p(A|B) = 0.362 + 0.138 - 0.036 = 0.464
p_a_or_b = (pb + pb - pa&b) ## formula for finding probability of A or B 
p_a_or_b ## result of the probability

```

Q4. What percent of the world population do not have a university degree and disagree with the statement 
   about men having more right to a job than women?

A4: The result is 0.536 equivalent to 53.6%
``` {r} 
## from the calculations done in the previous question we know that p(A|B) = 0.464, therefore
## p(!AB) = 1 - 0.464 = 0.536
p_not_a_b = 1 - (p_a_or_b) ## finding probability of NOT A & B 
p_not_a_b ## results
p_not_a_b*100 ## result in percentage

```
Q5. Does it appear that the event that someone agrees with the statement is independent of the event that they have a university degree or higher?

A5: The events are not independent considering the numeric results that come when we compare P(A&B) with P(A)^P(B) which are not equal, P(A&B)!=P(A)*P(B).
  
Q6. What is the probability that at least 1 in 5 randomly selected people to agree with the statement about men having more right to a job than women?

A6: The probability that at least 1 in 5 randomly selected agrees is 0.894 or 89.4%
```{r}
p_a <- pa # agree 36.2 = 0.362
p_na <- 1 - pa # don't agree 1 - 0.362 = 0.638
total = 1 - p_na^5 ## 1 in 5 randomly selected
## total = 1 - 0.638^5 = 1 - 0.10570691 = 0.89429308615
```

II. Data: As of 2009, Swaziland had the highest HIV prevalence in the world. 25.9% of this countrys population is infected with HIV. 
The ELISA test is one of the first and most accurate tests for HIV. For those who carry HIV, the ELISA test is 99.7% accurate. 
For those who do not carry HIV, the test is 92.6% accurate. If an individual from Swaziland has tested positive, 
what is the probability that he carries HIV? Create a tree diagram to calculate the probability.

Q1. If an individual from Swaziland has tested positive, what is the probability that he carries HIV?

P(HIV) = 0.259
P(!HIV) = 0.741
P(HIV_POS_E) = 0.997 -> ! = 0.003
P(HIV_NEG_E) = 0.926 -> ! = 0.038

A3: P(HIV_POSITIVE) = (0.259 * 0.997) + (0.926 * 0.038) = 0.258 + 0.035 = 0.293
    P(HIV_POS_TOTAL) = 0.258 / 0.293 = 0.8813
    88.13%
 
Q2. According to a 2013 Gallup poll, worldwide only 13% of employees are engaged at work (psychologically committed to their jobs and likely to be making positive contributions to their organizations). Among a random sample of 10 employees, what is the probability that 8 of them are engaged at work?

A3: dbinom(8, size = 10, p = 0.13)
    2.77842e-06
  
Q3. Recent study: “Facebook users get more than they give”

  friend requests: 40% made, 63% received at least one
  likes: liked 14 times, had their content “liked” 20 times, on average
  messages: sent 9 messages, received 12, on average
  tags:12% tagged a friend in a photo, but 35% tagged other findings:
  25% considered power users
  average Facebook user has 245 friends
  P(70 or more power user friends) = ?
  
A3:  0.112763
     sum(dbinom(70:245, 245, 0.25))


Q4. According to a 2014 Gallup poll, 56% of uninsured Americans who plan to get health insurance say they will do so through a government health insurance exchange. What is the probability that in a random sample of 10 people exactly 6 plan to get health insurance through a government health insurance exchange?

A4: 0.2427494
    dbinom(6, size = 10, p = 0.56)
  
  